Solutions to Integer Programming from Quantum Annealing

### 2020-08-26 paper/main.tex
update paper

\documentclass[prd,twocolumn,tightenlines,preprintnumbers,showpacs,superscriptaddress,notitlepage,nofootinbib,eqsecnum,floatfix,longbibliography]{revtex4}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb}
\usepackage{bm,bbm}
\usepackage{graphicx}
\usepackage{color}
\usepackage{dsfont}

@@

Formally, ILP is defined as the optimization of a linear function
{\color{blue}
\begin{align}
f(x) = &\min\limits_{x}(\sum_i c_i x_i),\\
\textrm{subject to} \quad & \sum_i A_{ai}x_i +b_a \leq 0 \\
& x_i \geq 0,\\
& x_i \in \mathbb{Z}
\end{align}
}
where $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations. The inequality constraints can be rewritten as an equality by introducing slack variables $s$ such that the ILP becomes
{\color{blue}
\begin{align}
f(x) = &\min(\sum_i c_i x_i),\\
\textrm{subject to} \quad & \sum_i A_{a i}x_i + s_a + b_a = 0,\\
& s_a \geq 0,\\
& s_a \in \mathbb{Z},\\
& x_i \geq 0,\\
& x_i \in \mathbb{Z}.
\end{align}
}
Since quantum annealing solves the QUBO problem, we extend the proposed algorithm to solve integer quadratic optimization problems such that
{\color{blue}
\begin{align}
f(x) = \min\limits_{x}(\sum_{ij} x_i d_{ij} x_j + \sum_i c_i x_i)
\end{align}
}
without the introduction of auxiliary qubits.

{\color{red} Paragraph on how this is solved classically. Include strengths and drawbacks of classical solution.}

@@

\begin{figure*}
	\centering
	\begin{tabular}{p{0.3\textwidth}p{0.3\textwidth}p{0.3\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/subaru_ds.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/subaru_mds4.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/subaru_mds2.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c}
	\end{tabular}
	\caption{Example of different dominating sets for $G(V, E)$. Vertices in the dominating set $D$ are highlighted in blue. {\bf{a)}} A dominating set of $G$ with domination number $\overline{\overline{D}} = 5$. {\bf{b)}} A minimal dominating set of $G$ with domination number of $\overline{\overline{D}} = 4$. We note that $D$ is also the maximum independent set of $G$. {\bf{c)}} The minimum dominating set of $G$ with domination number of $\overline{\overline{D}} = 2$.}
	\label{fig:dominating_sets}
\end{figure*}

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS.
We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.
\begin{figure*}
	\centering
	\begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
	\end{tabular}
	\caption{
		Example of greedy MDS algorithm:
		The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
		Connected nodes are neighbors of the DS (turquoise) or already in the DS.
		In (a), one finds three unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
		In the first step $a\to b$, randomly the node (4) is picked.
		In (b), one has two unconnected choices [(1), (2)] with each adjacent to another unconnected node.
		After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
		The found solution is a (minimal) dominating set but not a MDS.
		To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
	}
	\label{fig:mds-greedy}
\end{figure*}

@@

The target Hamiltonian can be written in the matrix form
\begin{widetext}
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
\mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] & - p(\mathbbm{1}+J^T)T^s\\
- p{T^s}^T(\mathbbm{1}+J)& p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + p \overline{\overline{V}}
\label{eq:matrix_form}
\end{align}
\end{widetext}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

@@

\subsection{Finding the Dominating Set with a Quantum Annealer}
\label{sec:discussion:qa}
\begin{figure}[b]
	\centering
		\includegraphics[width=0.5\columnwidth]{./figures/linear_crop.pdf}
	\caption{Linear graphs $G(n)$ used in this study. Nodes denote vertices of the graphs and lines are undirected edges.}
	\label{fig:linear}
\end{figure}

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=\columnwidth]{./figures/anneal_time_scaling.pdf}
	\caption{Probability of finding the MDS for $G(6)$ as a function of total annealing time.}
	\label{fig:at_scale}
\end{figure}

For the following studies, we perform experiments on the \texttt{DW\_2000Q\_5} solver. The annealing time is optimized to 600$\mu$s after performing a study on the $G(6)$ graph as shown in Fig.~\ref{fig:at_scale}. In Fig.~\ref{fig:baseline}, we show results from the baseline experiment, without modification to the default DWave annealing schedule, and observe improvement over random guessing. However, the experiment results reflect the same degeneracy structure as random guessing, and scale exponentially poorly for larger number of vertices $n$. We explore one avenue towards improving the experiment results by introducing per-qubit annealing offsets into the time evolution.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./figures/scaling_baseline.pdf}
	\caption{Baseline result of DWave (blue) compared to random guessing (red). The jagged nature of random guessing reflects the degeneracy of the ground state.}
	\label{fig:baseline}
\end{figure}

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=\columnwidth]{./figures/anneal_schedule.pdf}
	\caption{asdf}
	\label{fig:anneal_schedule}
\end{figure}

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=\columnwidth]{./figures/scaling_comparison_all.pdf}
	\caption{The legend labels the total range of the symmetric offset employed. A positive range indicates that the group of qubits with small values of $|h_i|$ are advanced (qubits with large values of $|h_i|$ are equivalently delayed), and vice versa for results labeled with a negative range.}
	\label{fig:dwave_offset}
\end{figure}

@@

\begin{figure}
	\centering
	%\includegraphics[width=\columnwidth]{./figures/dwave1us.pdf}
	\includegraphics[width=\columnwidth]{./figures/sim_deco.pdf}
	\caption{ {\color{red} New figure, Dwave on top of simulation (black).} The probability of finding the MDS for $G(2)$ from DWave. The machine parameters are identical to Sec.~\ref{sec:discussion:qa} with the exception of shortening the total annealing time to 1 $\mu s$ to accommodate reasonable simulation times. (Middle) Result from simulation with decoherence and (Bottom) simulation without decoherence. In both simulations, at +10\% offset, the probability of finding the ground state vanishes. The red simulation points are not present in the DWave result.}
	\label{fig:dwave1us}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./figures/spectrum.pdf}
	\includegraphics[width=\columnwidth]{./figures/spectrum_offset.pdf}
	\caption{{\color{red} Maybe combine these 2 plots to 1 as well? same color for same offset.} The energy spectrum of the first 5 states of the time-dependent Hamiltonian (Top) without annealing offsets and (Bottom) -10\% annealing offset. The spectrum in dimensionless units at the end of the anneal is an integer spectrum and correspond to the number of vertices in the dominating set. For $G(2)$, the MDS has 1 vertex. Annealing offsets lift the ground-state degeneracy.}
	\label{fig:spectrum}
\end{figure}

@@

\subsection{Many-Body Localization}

We provide further evidence that suggests the advancement of strong external fields leads to a delay in the phase transition, and increase in many-body localization, and results in a higher probability to recover the true ground state in this section.

@@

\begin{figure*}
	\centering
	\includegraphics[width=\columnwidth]{./figures/full_prob_deco.pdf}
	\includegraphics[width=\columnwidth]{./figures/hybridization.pdf}
	\caption{(Left) The time-dependent probability of resolving the ground state and (Right) the corresponding mutual information. (Top) Simulation with decoherence included and (Bottom) without decoherence.}
	\label{fig:prob_mi}
\end{figure*}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./figures/final_state_distribution.pdf}
	\caption{{\color{red}I need to add dwave final state distribution here too.}}
	\label{fig:final_state_distribution}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./figures/coherence.pdf}
	\caption{asdf}
	\label{fig:prob_mi}
\end{figure}

>>>>

\documentclass[prd,twocolumn,tightenlines,preprintnumbers,showpacs,superscriptaddress,notitlepage,nofootinbib,eqsecnum,
floatfix,longbibliography]{revtex4-2}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{bm,bbm}
\usepackage{graphicx}
\usepackage{xcolor}

@@

Formally, ILP is defined as the optimization of a linear function
\begin{align}
f(x) = &\min\limits_{x}(\sum_i c_i x_i),\\
\textrm{subject to } &\left \{
\begin{matrix*}[l]
	&\sum_i A_{ai}x_i +b_a \leq 0 \\
	&x_i \geq 0,\\
	&x_i \in \mathbb{Z}
\end{matrix*} \right.
\end{align}
where $i=1, \cdots,  N$ is the number of dependent variables and $a=1, \cdots, M$ the number of constraint equations. The inequality constraints can be rewritten as an equality by introducing slack variables $s$ such that the ILP becomes
\begin{align}
f(x) = &\min(\sum_i c_i x_i),\\
\textrm{subject to } &\left \{
\begin{matrix*}[l]
	&\sum_i A_{a i}x_i + s_a + b_a = 0,\\
	&s_a \geq 0,\\
	&s_a \in \mathbb{Z},\\
	&x_i \geq 0,\\
	&x_i \in \mathbb{Z}.
\end{matrix*} \right.
\end{align}
Since quantum annealing solves the QUBO problem, we extend the proposed algorithm to solve integer quadratic optimization problems such that
\begin{align}
f(x) = \min\limits_{x}\left(\sum_{ij} x_i d_{ij} x_j + \sum_i c_i x_i\right)
\end{align}
without the introduction of auxiliary qubits.

@@

\begin{figure*}
	\centering
	\begin{tabular}{p{0.2\textwidth}p{0.1\textwidth}p{0.2\textwidth}p{0.1\textwidth}p{0.2\textwidth}}
	\includegraphics[width=0.2\textwidth]{./new_figures/MDS_mds0.pdf}
&&
	\includegraphics[width=0.2\textwidth]{./new_figures/MDS_mds1.pdf}
&&
	\includegraphics[width=0.2\textwidth]{./new_figures/MDS_mds2.pdf}\\
	\centering\bf{a} && \centering\bf{b} && \centering\bf{c}
	\end{tabular}
	\caption{Example of different dominating sets for $G(V, E)$. Vertices in the dominating set $D$ are highlighted in blue. {\bf{a)}} A dominating set of $G$ with domination number $\overline{\overline{D}} = 2$. {\bf{b)}} A minimal dominating set of $G$ with domination number of $\overline{\overline{D}} = 2$. {\bf{c)}} The minimum dominating set of $G$ with domination number of $\overline{\overline{D}} = 1$.}
	\label{fig:dominating_sets}
\end{figure*}

For general graphs, existing "classical" algorithms either find minimal solutions in exponential time $\sim O( 1.5^n)$ \cite{Fomin2009, vanRooij2009} or approximate solutions in polynomial time. For example, greedy algorithms locally optimize decisions about which nodes to add to the dominating set.
Thus one is guaranteed to find a dominating set but not necessarily a MDS. {\color{red} Write some pseudo code for greedy search.}
%We demonstrate an example greedy algorithm in figure Fig.~\ref{fig:mds-greedy}.

\iffalse
\begin{figure*}
	\centering
	\begin{tabular}{p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}p{0.22\textwidth}}
	\includegraphics[width=0.20\textwidth]{./figures/greedy-1.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-2.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-3.pdf}
&
	\includegraphics[width=0.20\textwidth]{./figures/greedy-4.pdf}\\
	\centering\bf{a} & \centering\bf{b} & \centering\bf{c} & \centering\bf{d}
	\end{tabular}
	\caption{
		Example of greedy MDS algorithm:
		The algorithm iteratively adds vertices to the dominating set (black nodes) such that newly added nodes are maximally unconnected (white nodes).
		Connected nodes are neighbors of the DS (turquoise) or already in the DS.
		In (a), one finds three unconnected nodes [(2), (4), (7)] which are adjacent to three unconnected nodes each.
		In the first step $a\to b$, randomly the node (4) is picked.
		In (b), one has two unconnected choices [(1), (2)] with each adjacent to another unconnected node.
		After randomly selecting (2) when going from $b \to c$, only one unconnected node is remaining (6).
		The found solution is a (minimal) dominating set but not a MDS.
		To find the MDS using this greedy algorithm, the first node choice must be either (2) or (5).
	}
	\label{fig:mds-greedy}
\end{figure*}
\fi

@@

The target Hamiltonian can be written in the matrix form
\begin{widetext}
\begin{align}
E = &
\begin{pmatrix}
\Psi^x & \Psi^s
\end{pmatrix}
\begin{pmatrix}
\mathbbm{1} + p\left[J^T J + J^T + J - 2 \mathrm{Diag}_{\psi^x}(|J|) - \mathbbm{1} \right] & - p(\mathbbm{1}+J^T)T^s\\
- p{T^s}^T(\mathbbm{1}+J)& p\left[{T^s}^T T^s + 2\mathrm{Diag}_{\psi^s}(|T^s|)\right]
\end{pmatrix}
\begin{pmatrix}
\Psi^x\\ \Psi^s
\end{pmatrix} + p \overline{\overline{V}}
\label{eq:matrix_form2}
\end{align}
\end{widetext}
where $ |M| \equiv \sum_{\nu} M_{\nu \mu}$.

@@

\subsection{Finding the Dominating Set with a Quantum Annealer}
\label{sec:discussion:qa}
\begin{figure}[b]
	\centering
	\begin{tabular}{cc}
	$G(n):$ &
	\raisebox{-.4\height}{\includegraphics[width=0.8\columnwidth]{./new_figures/linear_graph.pdf}}
	\end{tabular}
	\caption{Linear graphs $G(n)$ used in this study. Nodes denote vertices of the graphs and lines are undirected edges.}
\label{fig:linear}
\end{figure}

@@

For the following studies, we perform experiments on the \texttt{DW\_2000Q\_6} solver. The annealing time is set to 500$\mu$s after performing a study on various annealing times the $G(6)$ graph. In Fig.~\ref{fig:baseline}, we show results from the baseline experiment, without modification to the default DWave annealing schedule, and observe improvement over random guessing. However, the experiment results reflect the same degeneracy structure as random guessing, and scale exponentially poorly for larger number of vertices $n$. We explore one avenue towards improving the experiment results by introducing per-qubit annealing offsets into the time evolution.

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./new_figures/DWave_scaling.pdf}
	\caption{Baseline result of DWave (black) compared to random guessing (blue). The jagged nature of random guessing reflects the degeneracy of the ground state. The legend labels the values of annealing offsets employed. A positive offsets (yellow) indicates that qubits with large values of $|h_i|$ are annealing with a delayed schedule, while negative offsets (green) delay the schedule of qubits with small values of $h_i$.}
	\label{fig:baseline}
\end{figure}

@@

\begin{figure}[b]
	\centering
	\includegraphics[width=\columnwidth]{./new_figures/anneal_schedule.pdf}
	\caption{asdf}
	\label{fig:anneal_schedule}
\end{figure}

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./new_figures/NN2_offset_scaling.pdf}
	\caption{ {\color{red} New figure, Dwave on top of simulation (black).} The probability of finding the MDS for $G(2)$ from DWave. The machine parameters are identical to Sec.~\ref{sec:discussion:qa} with the exception of shortening the total annealing time to 1 $\mu s$ to accommodate reasonable simulation times. (Middle) Result from simulation with decoherence and (Bottom) simulation without decoherence. In both simulations, at +10\% offset, the probability of finding the ground state vanishes. The red simulation points are not present in the DWave result.}
	\label{fig:dwave1us}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./figures/spectrum_all.pdf}
	%\includegraphics[width=\columnwidth]{./figures/spectrum_offset.pdf}
	\caption{{\color{red} NEED NEW PLOTS. Maybe combine these 2 plots to 1 as well? same color for same offset.} The energy spectrum of the first 5 states of the time-dependent Hamiltonian (Top) without annealing offsets and (Bottom) -10\% annealing offset. The spectrum in dimensionless units at the end of the anneal is an integer spectrum and correspond to the number of vertices in the dominating set. For $G(2)$, the MDS has 1 vertex. Annealing offsets lift the ground-state degeneracy.}
	\label{fig:spectrum}
\end{figure}

@@

\subsection{Many-Body Localization}
{\color{red} Probably need to rerwite this...}
We provide further evidence that suggests the advancement of strong external fields leads to a delay in the phase transition, and increase in many-body localization, and results in a higher probability to recover the true ground state in this section.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./new_figures/time_dependent_probability.pdf}
	\caption{(Left) The time-dependent probability of resolving the ground state and (Right) the corresponding mutual information. (Top) Simulation with decoherence included and (Bottom) without decoherence.}
	\label{fig:td_prob}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./new_figures/final_state_distribution.pdf}
	\caption{{\color{red}I need to add dwave final state distribution here too.}}
	\label{fig:final_state_distribution}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{./figures/coherence.pdf}
	\caption{asdf}
	\label{fig:prob_mi2}
\end{figure}
